Hydra Battles and Cie

### 2021-01-20 doc/chapter-primrec.tex
redaction of Ackermann chapter (continued)

This chapter contains some comments on Russel's library, as well as a few extensions.

@@

\section{Two challenges}

The following projects may require to look at other libraries, and or prove new lemmas.
Please send us your solutions (even if they are partial).

\index{Projects}

\begin{project}
Write a simple and readable proof that the Fibonacci function is primitive recursive.

We give the tail-recursive definition from~\cite{BC04}, chapter 9. See also
\href{https://coq-community.org/coq-art/ch9_function_specification/SRC/fib_tail.v}{this file}.
\begin{Coqsrc}

Fixpoint general_fib (a0 a1 n:nat) {struct n} : nat :=
  match n with
  | O => a0
  | S p => general_fib a1 (a0 + a1) p
  end.

Definition fib_tail (n:nat) := general_fib 1 1 n.
\end{Coqsrc}
\end{project}

\begin{project}
Prove formally that the Ackermann function is \emph{not} primitive recursive.

You may find a definition and some useful lemmas in
 \href{../theories/ordinals/Prelude/Iterates.html}%
{ordinals.Prelude.Iterates}.

>>>>

This chapter contains some comments on Russel's library, as well as a few extensions.
Contributions (under the form of comments, new examples or exercises) are welcomed!.

@@

\index{Exercises}

\begin{exercise}
Write a simple and readable proof that the Fibonacci function is primitive recursive.

\begin{Coqsrc}
Fixpoint fib (n:nat) : nat :=
  match n with
  | 0 => 1
  | 1 => 1
  | S ((S p) as q) => fib q + fib p
  end.
\end{Coqsrc}

\textbf{Hint:}  You may use as a helper the function which computes the pair 
$(\texttt{fib}(n+1),\texttt{fib}(n))$. 
Library \href{../theories/html/hydras.Ackermann.cPair.html}{Ackermann.cPair} contains
the definition of the encoding of $\mathbb{N}^2$ into $\mathbb{N}$, and the proofs that 
the associated constructor and projections are primitive recursive.

\end{exercise}

\index{Projects}
\begin{project}
Prove formally that the Ackermann function is \emph{not} primitive recursive.
You may find a definition of the Ackermann function and a few  lemmas in
 \href{../theories/ordinals/Prelude/Iterates.html}%
{ordinals.Prelude.Iterates}.
